---
title: "Catchment Characteristics explaining TOC"
subtitle: "An overview of the catchemnt characteristics and initial statistics run for explaing TOC concentrations in the SMED brownification project " 
author: "Anna Lackner"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  pdf_document:
    toc: true
    toc_depth: 3
header-includes: \usepackage{float} \floatplacement{figure}{H}
fontsize: 12pt
---

```{r, echo = F, results = FALSE, include = F}
knitr::opts_chunk$set(echo = F)
knitr::opts_chunk$set(message = F)
knitr::opts_chunk$set(warning = F)
knitr::opts_chunk$set(fig.pos = '!ht')
#knitr::opts_chunk$set(fig.pos = "H", out.extra = "")

# to make sure R can deal with the å and other weird swedish characters
Sys.setlocale("LC_ALL", "en_US.UTF-8")

set.seed(12)
```

\newpage

# Get Data

```{r}
library(tidyverse)
# loading the catchment data #fill in your own path to nmd data here
landuse <- read.csv("../input/catchment_characteristics/PLC8_landuse.csv")
```

```{r}
# load data and calculate response variable
# here the response variable will be the median org charge/mg of C for any stations that has more than 10 samples 
data <- read.csv("../results/chemistry/complete.csv") %>% filter(abs(charge_diff) < 0.5) %>% select(mvm_id, org_charge_eq_mg_C) %>% group_by(mvm_id) %>%
  summarise(row_count = n(), org_charge_eq_mg_C = median(org_charge_eq_mg_C)) %>% filter (row_count > 10) %>% select(-row_count)

# merge response and landuse as one dataframe called all

all <- data %>% left_join(landuse, by = join_by(mvm_id))

```


# OPLS

I chose to use a OPLS instead of a PLS. Orthogonal Partial Least Squares (OPLS) enables to separately model the variation correlated (predictive) to the factor of interest and the uncorrelated (orthogonal) variation. While performing similarly to PLS, OPLS facilitates easier interpretation. I used the R package [ropls](https://rdrr.io/bioc/ropls/man/opls.html) to perform the below analysis, its the underlying code of the SIMCA software as used in analysis such as [Larson et al., 2020](https://hess.copernicus.org/articles/26/4837/2022/).



```{r}
library(plotly)

# Define the function
plot_loading <- function(opls_model, response_label = "TOC", vip.threshold = 1.0, text_size = 12) {
  
  vip_data <- enframe(getVipVn(opls_model), name = "Variable", value = "VIP.score")
  
  # Extract predictive and orthogonal loadings
  loadings_ortho <- getLoadingMN(opls_model, orthoL = TRUE)
  loadings_pred <- getLoadingMN(opls_model, orthoL = FALSE)
  
  # Convert the loadings to a data frame
  loading_df <- data.frame(
    Variable = rownames(loadings_pred), 
    Predictive_Loading = loadings_pred[, 1],    # Extract the first column for predictive loading
    Orthogonal_Loading = loadings_ortho[, 1]    # Extract the first column for orthogonal loading
  ) %>% 
    left_join(vip_data, by = "Variable") %>%    # Join with VIP data on variable name
    mutate(color = ifelse(VIP.score < vip.threshold, "black", "blue"))
  
  # Filter for high VIP score variables (VIP > 1)
  loading_VIP <- loading_df %>% filter(VIP.score > 1)
  
  # Extract the response variable position
  x_1 <- opls_model@cMN[1, 1]  # Assuming the first coefficient as x
  y_1 <- 0.0                    # y set to 0 for the response variable
  
  # Create the interactive Plotly plot
  plot <- plot_ly(
    data = loading_df,
    colors = 'Paired',
    x = ~Predictive_Loading,
    y = ~Orthogonal_Loading,
    type = 'scatter',
    mode = 'markers',
    marker = list(
      size = ~VIP.score, 
      sizeref = 0.01,
      sizemode = 'area',
      color = ~color, 
      line = list(width = 0)
    ),
    text = ~Variable,
    hovertemplate = paste(
      "<b>%{text}</b><br><br>",
      "%{yaxis.title.text}: %{y}<br>",
      "%{xaxis.title.text}: %{x}<br>",
      "VIP score: %{marker.size:,}",
      "<extra></extra>"
    )
  ) %>%
    layout(
      title = "",
      xaxis = list(title = "Predictive Loading"),
      yaxis = list(title = "Orthogonal Loading"), 
      showlegend = FALSE
    ) %>% 
    # Add variable annotations for high VIP scores
    add_annotations(
      x = loading_VIP$Predictive_Loading,
      y = loading_VIP$Orthogonal_Loading - 0.03,
      text = loading_VIP$Variable,
      xref = "x",
      yref = "y",
      showarrow = FALSE,
      font = (list(size = text_size))
    ) %>% 
    # Add point for the response variable
    add_trace(
      x = x_1,
      y = y_1,
      mode = 'markers+text',
      type = 'scatter',
      marker = list(size = 20, color = 'red'),
      text = response_label,
      textfont = (list(size = text_size)),
      textposition = "top",
      hovertemplate = paste(
        "<b>", response_label, "</b><br><br>",
        "Response Variable: %{x}<br>"
      )
    ) 
  
  return(plot)
}
```


```{r}
library(ropls)

opls.model <- opls(all %>% select(-org_charge_eq_mg_C), all %>% select(org_charge_eq_mg_C), orthoI = 1, predI = 1)
```

```{r}
plot_loading <- function(opls.model, response_label = "charge density", vip.threshold = 1.0, text_size = 12)
```
