---
title: "Charge_density calculations"
author: Anna Lackner
date: today
format:
    html:
        embed-resources: true
        toc: true
        toc-depth: 3
        warning: false
        smooth-scroll: true
        code-fold: true
execute: 
  cache: true
---

# Overview
Based on 35 years of monitoring data of the water chemistry lab at the SLU Department of Aquatic Sciences and Assessment, the charge density of organic matter was calculated for over 40 000 samples across Swedish from water courses. 

## Methothodology
The following steps were performed to get final values of charge density (charge per mg of carbon). 

1.  Raw water chemistry downloaded from the MVM database, for all streams in the database that had at least 10 years of continuous monthly sampling of TOC in the period 1990-2023. 
2.  Preprocessing of water chemistry data to aggregate different measurnment methods for the same variable into single columns. Here samples which did not have enough other parameters measured that were needed in the following steps were dropped from the analysis. Which parameters were deemed necessary is further elaborated in Section Visual Minteq. 
3. Modelling of DOC charge density in Visual Minteq by sweeping through the ADOM/DOC parameter for each sample, calculating charge difference for each ADOM/DOC. 
4. Processing of Visual Minteq output to find the ADOM/DOC that best models the chemical equiblirum for each sample with the smallest charge difference and use it to calculate the charge density of each sample.
5. Collection of catchment characteristics of the stations. 
6. Statistical Analysis looking at both spatial and temporal variatioon of DOC charge density across the dataset.  

# Data source

- Downloaded 142 000 samples from [MVM Database](https://miljodata.slu.se/MVM/)
- Stations were preselected for streams that had TOC sampled 10 times a year for at least 10 consecutive years since 1990. 
- Stations with more than 5% Urban area were excluded from the downloading. 
- Data for 316 stations were downloaded. 

```{python map of stations}
import pandas as pd
import geopandas as gpd
import pandas as pd
import matplotlib.pyplot as plt
from shapely.geometry import Point

# Path to the CSV file
file_path = "../input/shapefiles/catch_316.zip"

# Load the CSV file into a DataFrame
gdf = gpd.read_file(f"zip://{sweden_shapefile}")
# Step 2: Convert to GeoDataFrame with SWEREF 99TM CRS
geometry = [Point(xy) for xy in zip(df["lat"], df["lon"])]
stations_gdf = gpd.GeoDataFrame(df, geometry=geometry, crs="EPSG:3006") 
stations_gdf.to_crs("EPSG:4326")  # SWEREF 99TM CRS

# Step 3: Load Sweden shapefile and ensure it's in SWEREF 99TM
sweden_shapefile = "../input/shapefiles/Sweden.zip"
sweden_gdf = gpd.read_file(f"zip://{sweden_shapefile}")
sweden_gdf = sweden_gdf.to_crs("EPSG:4326")  # Ensure CRS matches SWEREF 99TM

# Create the Plotly map
fig = go.Figure()

# Add Sweden polygon layer
fig.add_trace(go.Choropleth(
    z=[1] * len(sweden_gdf),
    locations=[i for i in range(len(sweden_gdf))],
    showscale=False,
    coloraxis="coloraxis",
    geojson=sweden_gdf.geometry.__geo_interface__,
    hoverinfo="skip",  # No hover for Sweden, just a base layer
))

# Add catchment areas with grey outlines
fig.add_trace(go.Choropleth(
    z=[1] * len(catch_gdf),
    locations=[i for i in range(len(catch_gdf))],
    showscale=False,
    coloraxis="coloraxis",
    geojson=catch_gdf.geometry.__geo_interface__,
    hoverinfo="skip",  # No hover for catchment areas
    line_color="black",  # Black outline for the catchments
    fillcolor="grey",  # Grey fill for the catchments
))

# Add station points with hover info showing station ID
fig.add_trace(go.Scattermapbox(
    lat=stations_gdf['lat'],
    lon=stations_gdf['lon'],
    mode='markers',
    marker=dict(
        size=10,
        color='red',
        opacity=0.8,
    ),
    text=stations_gdf['stationID'],  # Hover text showing station ID
    hoverinfo='text',  # Display only station ID on hover
))

# Update layout for the map
fig.update_layout(
    title="Catchment Areas and Stations",
    mapbox=dict(
        style="open-street-map",  # Base map style
        center=dict(lat=58, lon=15),  # Approximate center of Sweden
        zoom=5,  # Zoom level
    ),
    coloraxis_showscale=False,  # Hide color scale since it's not needed
)

# Show the map
fig.show()
```

# Preprocessing

- 42 907 samples across 187 stations
- Median number of samples per station: 208

```{r}
#| echo: false
#| fig-width: 5 
#| fig-height: 4
#| fig-cap: Histogram of number of samples per station used in further analysis. There are 42 907 samples across 187 stations. In further analysis only stations with more than 5 samples were considered, reducing the number of stations to 173.
#| fig-alt: Histogram of number of samples per station
#| label: fig-hist_stations 

library(tidyverse)

data <- read.csv("../results/chemistry/complete.csv")
mvm_ids <- unique(data$mvm_id)
stations <- read.csv("../input/catchment_characteristics/catch_landuse(NMD).csv") %>% select(mvm_id, stationCoordinateX, stationCoordinateY, area_ARO_m2)

# Step 1: Count rows for each mvm_id
mvm_counts <- data %>%
  group_by(mvm_id) %>%
  summarise(sample_count = n())

# Step 2: Plot histogram of row counts
ggplot(mvm_counts, aes(x = sample_count)) +
  geom_histogram(binwidth = 12, fill = "skyblue", color = "black", alpha = 0.7) +
  labs(title = "",
       x = "Samples per station",
       y = "Frequency") +
  theme_minimal()

data %>% select(mvm_id) %>%
  group_by(mvm_id) %>%
  arrange(mvm_id) %>%       # Ensure data is ordered within each mvm_id (if needed)
  slice(1) %>%             # take the first row of each group
  ungroup() %>% left_join(stations) %>% left_join(mvm_counts) %>% filter(sample_count >= 5) -> stations

mvm_ids <- stations$mvm_id

data %>% filter(mvm_id %in% mvm_ids) -> data
```

# Visual Minteq

## Required parameters

- silicon set to same value as in Sjösted et al, 2010.
- alkalinity/acidity was added as Na and Cl concentration respectively. 
- bold parameters had to be measured parameters in the raw data, other parameters were set to 0 were not measured.

| Name        | Var           | Valance |
|-------------|---------------|---------|
| **Aluminium**   | Al_mol        |         |
| Copper      | Cu_mol        | +3      |
| Manganese   | Mn_mol        | +2      |
| **TOC**         | TOC_mol       | -x      |
| Zinc        | Zn_mol        |         |
| **SO4**         | SO4_mol       | -2      |
| **NO3**         | NO3_N_mol     | -1      |
| **Cl**          | Cl_alk_mol    | -1      |
| **Pottasium**   | K_mol         | +1      |
| **Calcium**     | Ca            | +2      |
| **Sodium**      | Na_acid_mol   | +1      |
| **Magnesium**   | Mg            | +2      |
| Iron        | Fe            | +3      |
| **Silicon**     | Si            |         |
| Ammonium    | NH4           | +1      |
| Fluorid     | F             | -1      |


## Modelling Set-up

- temperature set to 10C
- pH 5.6
- Ferrihydrite as possible solid phase
- AlOH3 as possible solid phase
- ADMO/DOC sweep from 0.05 to 3.5

# Post Processing

For each sample the ADOM/DOC was selected that had the minimum absolute value of charge difference. Charge difference was calculated according to Visual Minteq as: 

$$
\text{Charge difference (\%)} = 100 \times \left| \frac{\text{SumA} - \text{SumC}}{\text{SumA} + \text{SumC}} \right|
$$

```{r}
#| fig-cap: Histogram of ADOM/DOC for samples before and after filtering for 173 stations (number of samples > 5).  
#| fig-alt: Histogram of ADOM/DOC.
#| label: fig-hist_adom_doc_after
#| fig-subcap:
#|    - Before filtering of charge difference < 0.5%
#|    - After filtering of charge difference < 0.5%

ggplot(data, aes(x = adom_doc)) +
  geom_histogram(binwidth = 0.05, fill = "skyblue", color = "black", alpha = 0.7) +
  labs(title = paste("Number of samples: ", nrow(data)),
       x = "ADOM/DOC",
       y = "Frequency") +
  theme_minimal()

ggplot(data %>% filter(abs(charge_diff) < 0.5), aes(x = adom_doc)) +
  geom_histogram(binwidth = 0.05, fill = "skyblue", color = "black", alpha = 0.7) +
  labs(title = paste("Number of samples: ", nrow(data %>% filter(abs(charge_diff) < 0.5))),
       x = "ADOM/DOC",
       y = "Frequency") +
  theme_minimal()
```

```{r}
#| fig-cap: Histogram of charge density for samples before and after filtering. 
#| fig-alt: Histogram of charge density.
#| label: fig-hist_cd_after
#| fig-subcap:
#|    - Before filtering of charge difference < 0.5%
#|    - After filtering of charge difference < 0.5%

ggplot(data, aes(x = org_charge_eq_mg_C)) +
  geom_histogram(fill = "skyblue", color = "black", alpha = 0.7) +
  labs(title = "",
       x = "charge density (charge/ mg C)",
       y = "Frequency") +
  theme_minimal()

ggplot(data %>% filter(abs(charge_diff) < 0.5), aes(x = org_charge_eq_mg_C)) +
  geom_histogram(fill = "skyblue", color = "black", alpha = 0.7) +
  labs(title = "",
       x = "charge density (charge/ mg C)",
       y = "Frequency") +
  theme_minimal()
```


# Catchment Characteristics

- [Soil Depth](https://www.sgu.se/en/products/maps/map-viewer/jordkartvisare/soil-depth/)
- [PLC8 NMD Land use data](https://helcom.fi/helcom-at-work/projects/plc-8/)
- [Discharge](https://www.smhi.se/data/hydrologi/vattenwebb/data-for-delavrinningsomraden-sotvatten-1.118236)
- [Climate](https://www.smhi.se/data/ladda-ner-data/griddade-nederbord-och-temperaturdata-pthbv)
- [Peat area](https://www.slu.se/en/environment/statistics-and-environmental-data/search-for-open-environmental-data/comprehensive-peat-map-of-the-forest-land/#downloadInfo)
- [NDVI](https://www.usgs.gov/landsat-missions/landsat-normalized-difference-vegetation-index)

```{r}
library(tidyverse)
soil_depth <- read.csv("../input/catchment_characteristics/soil_depth.csv") %>% mutate(soil_depth_iqr = X75th_percentile - X25th_percentile, soil_depth_mean = mean) %>% select(mvm_id, soil_depth_iqr, soil_depth_mean) %>% filter(mvm_id %in% mvm_ids)%>% mutate(mvm_id = as.factor(mvm_id))
```

```{r}
landuse <- read.csv("../input/catchment_characteristics/PLC8_landuse.csv") %>% filter(mvm_id %in% mvm_ids) %>% mutate(mvm_id = as.factor(mvm_id))
```

```{r}
#| fig-cap: Histogram of climate variables.
#| label: fig-climate 
#| fig-subcap:
#|    - mean annual temperature
#|    - total annual precipitation
#| fig-alt: Histogram of charge density.

climate <- read.csv("../input/climate/daily_climate.csv") %>% filter(mvm_id %in% mvm_ids) %>% mutate(date = as.Date(time), year = year(date)) %>% filter (year >= 1990)  %>% group_by(mvm_id, year) %>% summarize(mean_annual_temp = mean(tas_avg), annual_precip= sum(pr_avg)) %>% ungroup() %>% group_by(mvm_id) %>% summarize(mean_annual_temp = mean(mean_annual_temp), annual_precip= mean(annual_precip)) %>% ungroup() %>% mutate(mvm_id = as.factor(mvm_id))

ggplot(climate, aes(x = mean_annual_temp)) +
  geom_histogram(fill = "skyblue", color = "black", alpha = 0.7) +
  labs(title = "",
       x = "mean annual temperature (C)",
       y = "Frequency") +
  theme_minimal()

ggplot(climate, aes(x = annual_precip)) +
  geom_histogram(fill = "skyblue", color = "black", alpha = 0.7) +
  labs(title = "",
       x = "total annual precipitation (mm)",
       y = "Frequency") +
  theme_minimal()
```

```{r}
#| fig-cap: Discharge for all catchments in the DOC dissociation study. 
#| fig-alt: Histogram of charge density.
#| label: fig-discharge
#| fig-subcap:
#|    - Boxplot of all the catchments together.
#|    - Example of median annual discharge for 4 different catchments.
#
discharge <- read.csv("../input/discharge/daily_discharge.csv") %>% filter(mvm_id %in% mvm_ids) %>% mutate(date = as.Date(Datum), year = year(date)) %>% group_by(mvm_id, year) %>% summarize(median_q_m3_s = median(q), p25_q_m3_s  = quantile(q, probs = 0.25, na.rm = TRUE) , p75_q_m3_s = quantile(q, probs = 0.75, na.rm = TRUE)) %>% ungroup() %>% left_join(stations)%>% mutate(mvm_id = as.factor(mvm_id)) 

ggplot(discharge, aes(group = year, x = year, y = log(median_q_m3_s))) +
geom_boxplot()

ggplot(discharge %>% filter (mvm_id %in% c(21, 180,33,222 )) %>% mutate(q_spec= (median_q_m3_s / area_ARO_m2)*1000), aes( x = year, y = q_spec, color = mvm_id)) +
geom_line()+
labs( y = "Specific discharge (mm/s)")

discharge %>% mutate(q_spec= (median_q_m3_s / area_ARO_m2)*1000, p25_spec = (p25_q_m3_s / area_ARO_m2)*1000, p75_spec = (p75_q_m3_s / area_ARO_m2)*1000 ) %>%
group_by(mvm_id) %>% summarise(
    mean_Q_spec = mean(q_spec), p25_Q_spec = mean(p25_spec), p75_Q_spec = mean(p75_spec), mean_Q = mean(median_q_m3_s)) %>% ungroup() -> discharge_y
```

```{r}
#| fig-cap: Summer NDVI June-August for all catchments in the DOC dissociation study. 
#| fig-alt: Histogram of charge density.
#| label: fig-ndvi
#| fig-subcap:
#|    - Boxplot of all the catchments together.
#|    - Example of summer NDVI for 4 different catchments.


ndvi <- read.csv("../input/catchment_characteristics/NDVI.csv")  %>% filter(mvm_id %in% mvm_ids) %>% mutate(date = make_date(year,month,15)) %>% filter (month %in% c(6,7,8)) %>% group_by(mvm_id, year) %>% summarize(summer_NDVI = mean(NDVI_median, na.rm = TRUE)) %>% ungroup() %>% mutate(mvm_id = as.factor(mvm_id)) 

ggplot(ndvi, aes(group = year, x = year, y = summer_NDVI)) +
geom_boxplot()

ggplot(ndvi %>% filter (mvm_id %in% c(21, 180,33,222 )), aes( x = year, y = summer_NDVI, color = mvm_id)) +
geom_line()

ndvi_y <- ndvi %>% group_by(mvm_id)%>% summarize(summer_NDVI = mean(summer_NDVI, na.rm = TRUE)) %>% ungroup()
```
```{r}
peat <- read.csv("../input/catchment_characteristics/peat_area.csv") %>% rename(water = X0, mineral_soil = X1, peat_a30 = X2, peat_a40 = X3, peat_a50 = X4) %>% mutate(peat = ((peat_a30 + peat_a40 + peat_a50) / sum), water = (water / sum), mineral_soil = (mineral_soil / sum )) %>% select(peat, mvm_id, water, mineral_soil)%>% filter(mvm_id %in% mvm_ids)  %>% mutate(mvm_id = as.factor(mvm_id))

```
## Summary
```{r}
# library(kableextra)
catch_char <- landuse %>% 
        left_join(peat %>% select(-water), by = join_by(mvm_id)) %>%
        left_join(soil_depth, by = join_by(mvm_id)) %>%
        left_join(climate, by = join_by(mvm_id))%>% 
        left_join(ndvi_y, by = join_by(mvm_id)) %>%
        left_join(discharge_y) %>% 
        left_join(stations %>% select(-sample_count ) %>% mutate(mvm_id = as_factor(mvm_id))) %>% 
        mutate(log_mean_Q = log(mean_Q)) %>% select(-mean_Q)

library(vtable)

sumtable(catch_char %>% select(-mvm_id), out ="kable", digits = 3)
```

# Temporal Patters

For each individual catchemnt split into three time periods, has to have at least 2 years samples in each time period for it to work. 
 

# Spatial Patterns

```{r}
# merge response and landuse as one dataframe called all
all <- data %>% filter(abs(charge_diff) < 0.5) %>%
 group_by(mvm_id) %>% summarise(org_charge_eq_mg_C = median(org_charge_eq_mg_C), mean_toc = median(TOC_mol)) %>%
 ungroup() %>% mutate(mvm_id = as_factor(mvm_id))%>% 
 select(mvm_id, org_charge_eq_mg_C, mean_toc) %>%
 left_join(catch_char, by = join_by(mvm_id))

all %>% write.csv("../results/r_py/pls_input.csv")
```

# OPLS

I chose to use a OPLS instead of a PLS for relating the results to catchment characteristics. Orthogonal Partial Least Squares (OPLS) enables to separately model the variation correlated (predictive) to the factor of interest and the uncorrelated (orthogonal) variation. While performing similarly to PLS, OPLS facilitates easier interpretation. I used the R package [ropls](https://rdrr.io/bioc/ropls/man/opls.html) to perform the below analysis, its the underlying code of the SIMCA software as used in analysis such as [Ehnvall et al., 2023](https://www.sciencedirect.com/science/article/pii/S0048969723037555).

## Organic charge

### OPLS
```{r}
source("../src/sourcecode.R")
library(ropls)
library(plotly)

opls.model <- opls(all %>% select(-c(org_charge_eq_mg_C, mvm_id)), all$org_charge_eq_mg_C, orthoI = 1, predI = 1)
```

```{r}
library(tibble)
source("../src/sourcecode.R")
plot_loading(opls.model, response_label = "Org. charge density", vip.threshold = 1.0, text_size = 12)
```

### PLS

Here a PLS was run instead of an OPLS for comparison
```{r}
pls.model <- opls(all %>% select(-c(org_charge_eq_mg_C, mvm_id)), all$org_charge_eq_mg_C)
```

## TOC
```{r}
# TOC
opls.model.TOC <- opls(all %>% select(-c(org_charge_eq_mg_C, mvm_id, mean_toc)), all$mean_toc, orthoI = 1, predI = 1)
```

```{r}
plot_loading(opls.model.TOC, response_label = "TOC", vip.threshold = 1.0, text_size = 12)
```

```{r}
pls.model <- opls(all %>% select(-c(org_charge_eq_mg_C,mean_toc, mvm_id)), all$mean_toc)
```

